Astrophysical calibration of gravitational-wave detectors

### 2015-07-02 docs/cosmic_cal.tex
Some major changes to the paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\usepackage {latexsym}
\usepackage {graphicx}
\usepackage{iopams}
\usepackage {color}

%\documentclass[12pt]{iopart}
%\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
%\usepackage {amsmath}
%\usepackage {amssymb}
%\usepackage {amsfonts}
%\usepackage {amsthm}
%\usepackage {mathrsfs}
%\usepackage {natbib}
%\usepackage {latexsym}
%\usepackage {graphicx}
%\usepackage {dsfont}
%\usepackage {times}
%\usepackage {txfonts}
%\usepackage {rotating}
%\usepackage {wasysym}
%\usepackage {multirow}
%\usepackage {hhline}
\usepackage {hyperref}
%\usepackage {color}
\usepackage {bm}
%\usepackage{appendix}
\usepackage{acronym}
%\usepackage{dcolumn}   % needed for some tables
\usepackage {url}
\usepackage[normalem]{ulem}   % temporal one in draft.

%\usepackage{harvard}

@@

\newcommand{\aap}{A\&A }
\newcommand{\apj}{ApJ  }
\newcommand{\apjl}{ApJL  }
\newcommand{\apjs}{ApJS  }
\newcommand{\prd}{Phys. Rev. D  }
\newcommand{\nat}{Nature}
\newcommand{\araa}{ARA\&A}
\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society  }
\begin{document}

@@

\author{L.~Wright$^1$, M.~Pitkin$^1$ \& C.~Messenger$^1$}
\address{$^1$ SUPA, School of Physics and Astronomy, University of
  Glasgow, Glasgow G12 8QQ, United Kingdom}
\ead{1002990W@student.gla.ac.uk}

\begin{abstract}
  We present an investigation into the potential of calibrating \MP{change to: ``assessing the validity of the calibration''?}
  gravitational wave detector outputs through the use of standard
  sirens. Such signals, as measured via gravitational wave
  observations provide an estimated luminosity distance which is
  subject to uncertainties in the calibration of the data.  If a host
  galaxy is identified for a given source then its redshift can be
  obtained and combined with current knowledge of the cosmologic
  parameters.  This will yield the true luminosity distance and allow
  the direct comparison with the estimated value.  Discrepancies can
  then be used to correct the original calibration.~\cm{Update at end}
\end{abstract}

% acronym definitions
\acrodef{GW}[GW]{gravitational wave}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{MWEG}[MWEG]{Milky Way equivelent galaxies}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{NSBH}[NSBH]{neutron star black hole}
\acrodef{SNR}[SNR]{signal-to-noise ratio}
\acrodef{GRB}[GRB]{gamma-ray burst}
\acrodef{sGRB}[sGRB]{short gamma-ray burst}
\acrodef{CBC}[CBC]{compact binary coalescence}
\acrodef{EM}[EM]{electro-magnetic}

@@

% introduce GWs and BNS systems
It is expected that the advanced generation of interferometric \ac{GW}
detectors will detect waves emitted from $O(10s)$ of compact binary
coalescences.  One such class of these cataclysmic events, the inspiral and
merger of \ac{BNS} systems will be detected out to a maximum range of $\approx
450$ Mpc. Assuming the current best estimates for the cosmological parameters,
this is equivelent to a redshift $z\approx 0.1$. As noted by \cite{1986Natur.323..310S} \ac{CBC}
systems can be used as cosmiological distance markers, otherwise known as
``standard sirens'' (analogous to the \ac{EM} standard candles). Standard sirens
allow ...

@@

\MP{The gravitational wave strain is measured through the differential arm length, $\Delta L$, changes
of the interferometer via $h = \Delta L(f) / L$, where $L$ is the total arm length.
Calibration is required to relate the actual measured interferometer error signal ouptut
$e(f)$ to $\Delta L$. This relation is known as the length response function, $R(f,t)$,
defined such that
\begin{equation}
\Delta L(f) = R(f,t) e(f),
\end{equation}
where $R$ varies slowly in time (in comparison to transient signal timescales).
Calculation of $R$ requires the measurement of various functions within a control feedback loop
(see \cite{Vitale:2012} for a brief overview of this), which are subject to
uncertainties. In this study we will assume an estimate of $R$ is available (although
in theory we could take on the role of estimating $R$ itself), but that it differs from the
truth through some unknown scale factor, $s$, so that
\begin{equation}
h(f) = s h_{\rm m}(f) = s \frac{R(f,t) e(f)}{L},
\end{equation}
where $h(f)$ is the true strain and $h_{\rm m}(f)$ is the measured strain.
In this analysis we will simplify the situation by assuming that $s$ is a constant with
respect to frequency, but in future studies that assumption could be dropped and $s$
could take some functional form, or piece-wise fit, with respect to $f$.
}

% Describe our model for the calibration
In this work the calibration correction factor used throughout is simply a scale
factor, $s$. This effects the data set collected by a simple multiplication:
%
\begin{equation}\label{eq:scaledata}
  h(t) = s\times h_{\mathrm{m}}(t)
\end{equation}
%
where $h_{\mathrm{m}}(t)$ is the measured detector strain, and
$h(t)$ is the exact perfectly calibrated strain. \sout{A description of the
collected data is given in Section~\ref{sec:cbc}}. The scale factor being
parameterised, $s$, represents all known sources of error in the system. This
is not entirely realistic as $s$ is known to be a function of frequency.
Testing a more realistic calibration error is work to be done in the
future.~\cm{Need to clean this up and expand the discussion}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A single event with GRB counterpart\label{sec:single}}

@@

The frequency-domain polarisation amplitudes are then defined as
%
\begin{eqnarray}
  \gws_{+}(f) &=& \frac{1+\cos^{2}\iota}{2D_{L}}
\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)}\\
  \gws_{\times}(f) &=&
 i\frac{\cos^{2}\iota}{D_{L}}\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)}
\end{eqnarray}
%
where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$,
and $\iota$ is the inclination angle.  The \ac{GW} strain measured
at the $k$'th detector is then given by
%
\begin{equation}
  \label{eq:gravsig}
   \gws^{k}(f) = \big[ F_{+}^{k}(\alpha, \delta, \psi)\gws_{+}(f) +
F_{\times}^D(\alpha, \delta, \psi)\gws_{\times}(f)\big]
\end{equation}
%
where $F_{+},F_{\times}$ are the antenna response functions which are dependent
upon the polarisation angle $\psi$ and the sky position of the source
$\alpha,\delta$.

@@

% Define the likelihood function we will be using
We consider that the measured data in any detector is defined as
%
\begin{equation}
  \tilde{d}_{k}(f)=\tilde{n}_{k}(f)+s\gws_{k}(f,\vec{\theta})
\end{equation}
%
where $\tilde{n}(f)$ is the Fourier transform of Gaussian noise drawn from a
distribution consistent with the power-spectral-density of the advanced
detector noise.

\subsection{Analysis method}

\MP{Our aim is to estimate the marginal posterior probability distribution function (pdf)
for the unknown calibration scale factor. We use Bayes' theorem for which we need to define
a likelihood function and prior pdfs on the signal parameters being estimated. In this
analysis we use a nested sampling algorithm \cite{Skilling}\footnote{The code used was based
on {\tt matlabmultinest} found at \url{https://github.com/mattpitkin/matlabmultinest}.}
to provide samples from the marginal posteriors that we require.}
We use a Gaussian likelihood function given by
%
%\begin{equation}
%  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta
%f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})}\Bigg]
%\end{equation}
\begin{equation}
  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm det}}
  \sum_{i=i_{f_{\rm low}}}^{i_{f_{\rm high}}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} \gws_{k,i}(\vec{\theta})
  - \frac{1}{2}\left(\gws^*_{k,i}(\vec{\theta})\gws_{k,i}(\vec{\theta}) + \tilde{d}^*_{k,i}\tilde{d}_{k,i}\right) \right\}}}{S_{k,i}}\right]}
\end{equation}
% 
where $\Delta f$ is the frequency resolution of the data, $S_k$ is the
$k^{\rm th}$ detector's one sided noise power spectral density, and the $i$ indices
increment over frequency with $i_{f_{\rm low}}$ and $i_{f_{\rm high}}$ corresponding to
the lower and upper range in frequencies used for our
analsysis. \MP{Need to say the
range of frequencies we integrate over, which I think was 40\,Hz to 1600\,Hz from looking
at Liam's code.} The source parameters, and the unknown scale factor, have been subsumed
into a vector $\vec{\theta}$.

@@

\subsection{Simulations}

% mention the no-noise issue
We note that the analysis performed here used zero-noise realisations in which
no actual noise was added to our signals.  The likelihood functions still
contain the correct \ac{PSD} and hence the accuracy with which the parameters
are recovered (indicated by the widths of the posterior distributions) is
representative of the statistical errors in realistic noise scenarios.

% Describe the ensemble of simulations used

For coalescing binary neutron stars of masses
$m_{1,2} = 1.4\,M_{\odot}$ and  $D_{L}200\,Mpc$, the expected SNR is roughly 8.
The SNR was determined from $D_{L} = [10,200]\,Mpc$  using a randomly chosen
sky position which defines the antenna patterns and the GW signal. The range of
luminosity distance used in this paper represented a high level of SNR up to
the distance at which the first predicted detectable GW source; $D_{L} =
200\,Mpc$. The detector limits for the advanced era is a luminosity distance of
$\sim 450\,Mpc$. \sout{This was not initially tested as the SNR was low enough to
cause problems with computational time. It was however tested in the random sky
position simulations, see section~\ref{sec:rndsp}. The initial value for the
calibration error used in the analysis was chosen randomly.}

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{DvsSNR-mult-det}
  \caption{The distance and SNR relationship for all three detectors:
LIGO Hanford, H1, LIGO Livingston, L1, and VIRGO.}
  \label{fig:distsnr}
\end{figure}

To begin with, parameter estimation was carried out on the simplest case: a
single source of GWs using a single detector, LIGO Hanford, at different levels
of SNR. The aim was to resolve the calibration scale factor with the highest
degree of certainty.  Before testing the uncertainty resolution using a the
full range of prior parameters, we fixed all values to known signal values and
only searched over the scale factor. Results shown for with and without a full
prior range are given in figures~\ref{fig:sd-empty-D}
and~\ref{fig:sd-non-empty-D} respectively.

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{SD_empty_D10_200}
  \caption{\textit{Posterior distribution on the calibration scale factor for
LIGO Hanford. Simulations carried out using an entirely known source of GWs.}}
  \label{fig:sd-empty-D}
\end{figure}

The figure above is a histogram of the calibration error, with each curve
representing a different distance (SNR). The black line at $s = 0.8$ is the
initial scale factor value used to multiply the unanalysed data set. This
representation of the initial calibration scale factor will be continuous
throughout the paper.  One would expect with lower SNR (greater $D_{L}$), the
estimation of the parameter would become less certain. This is clearly shown in
figure~\ref{fig:sd-empty-D} as the standard deviations of the curves increases
with smaller SNR/larger luminosity distances. For this entirely known source of
GWs, we would expect to find the uncertainty in $s$ to be approximately
$1/$SNR; this is what we see in figure~\ref{fig:sd-empty-D}. The error in
estimating $s$ for $D_{L} = 200 \,Mpc$ is roughly $\sim 20\%$. This is larger
than the current method of obtaining the uncertainty in calibration. A full
test of the new method proposed is carried out in section~\ref{sec:rndsp}.

The next estimation of the calibration scale factor was performed with a full
range of prior parameters, giving a more realistic system. As before, the
simulations were carried out for differing levels of SNR and a histogram of the
calibration error was made, seen in figure~\ref{fig:sd-non-empty-D}.

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{SD_non_empty_D10_200}
  \caption{\textit{Posterior distribution on the calibration scale factor for
LIGO Hanford. Simulations carried out with some parameters in the prior.}}
  \label{fig:sd-non-empty-D}
\end{figure}

The error in the calibration scale factor, as before, increases with decreasing
SNR (increasing $D_{L}$). The magnitude of this error however is larger than in
the previous simulation. With a full prior range given in
section~\ref{sec:pandl}, the standard deviation increases; at $D_{L} =
200\,Mpc$ the error in $s$ is $\sim 24\%$. When comparing this new method of
calibration scale factor estimation with the current, feedback loop system
method used today, this method is not as accurate. However, it should again be
noted that this method is an independent estimate of the calibration accuracy.

\sout{A network of detectors are used throughout GW research. Being able to show the
effect of the calibration error and recovering its uncertainty in the analysis
would help in giving a detection more confidence. In this section, the
simulations were carried out using multiple detectors; LIGO Hanford, LIGO
Livingston, and Virgo. In reality, multiple detectors are used in order to
determine the (rough) sky position of a signal. Due to their locations, if one
site detects a signal before the others, then the position of the source can be
roughly worked out. The aim was, as before, to estimate the calibration scale
factor with the least amount of uncertainty for differing levels of SNR for a
single source of GWs.}

For a network of detectors, the overall likelihood function becomes a product
of each likelihood for all detectors, i.e.

\begin{equation}
  \label{eq:mult-likeli}
  p(\vec{d}| \vec{\theta}, \curlH) = \prod \limits_D p(\vec{d}^D| \vec{\theta},
\curlH)
\end{equation}

where $D = \{H,L,V\}$. The initial calibration scale factors where again chosen
at random ($s = [4.6, 2.5, 6.1]$ for LIGO Hanford, LIGO Livingston, and Virgo
respectively). The simulations were the same as before, using the same
parameter values. The GRB assumption was again tested in these simulations as a
handful of parameters were relaxed and given some uncertainty. The results of
these simulations are shown in figure~\ref{fig:mult-empty-D} and
figure~\ref{fig:mult-non-D}.

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{MD_empty_D10_200}
  \caption{Posterior distribution on the calibration scale factor for
all three detectors. Simulations carried out using an entirely known source of
GWs.}
    \label{fig:mult-empty-D}
\end{figure}

Figure~\ref{fig:mult-empty-D} shows the histograms of the calibration scale
factor for each detector. These simulations used an entirely known source of
GWs, at a randomly chosen sky position (the same sky position as before). As
the figure shows both LIGO sites, Hanford and Livingston, are able to resolve
the error with a measurable degree of uncertainty. The standard deviation for
$s$ at $D_{L}=200\,Mpc$ is $24\%$ and $41\%$ of the true scale factor for
Hanford and Livingston respectively. The estimation for the calibration error
using VIRGO however, was not as accurate. It can be seen that, for increasing
luminosity distances, the resolution of $s$ becomes increasingly uncertain
until the point at which the distribution becomes almost uniform; $D_{L}
\approx 120\,Mpc$. There are a few reasons why this may be. Firstly, the sky
position used may not be 'ideal' for generating the antenna patterns. The most
probable cause of the estimation being very poor will lie in the detector
itself. The interferometer itself is not as sensitive the two LIGO sites, hence
the estimation of parameters using this site will not be as precise.

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{MD_non_empty_D10_200}
  \caption{Posterior distribution on the calibration scale factor for
all three detectors. Simulations carried out with some prior parameters.}
  \label{fig:mult-non-D}
\end{figure}

The histograms shown in figure~\ref{fig:mult-non-D} were made using the full
range of prior parameters. The uncertainty in the calibration error is larger
with a less known source of GWs. VIRGO, again, has the least certain
calibration error resolution with the distribution becoming even more uniform.
The standard deviation for both LIGO curves become larger with increasing
luminosity distance as before. For $D_{L} = 200Mpc$,  the standard deviation is
$\tilde 30\%$ of the true value. When comparing this to the current method,
this again is not as accurate.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[10pt]{iopart}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{iopams}
\usepackage{color}

%\documentclass[12pt]{iopart}
%\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
%\usepackage {amsmath}
%\usepackage {amssymb}
%\usepackage {amsfonts}
%\usepackage {amsthm}
%\usepackage {mathrsfs}
%\usepackage {natbib}
%\usepackage {latexsym}
%\usepackage {graphicx}
%\usepackage {dsfont}
%\usepackage {times}
%\usepackage {txfonts}
%\usepackage {rotating}
%\usepackage {wasysym}
%\usepackage {multirow}
%\usepackage {hhline}
\usepackage{hyperref}
%\usepackage {color}
\usepackage{bm}
%\usepackage{appendix}
\usepackage{acronym}
%\usepackage{dcolumn}   % needed for some tables
\usepackage{url}
\usepackage[normalem]{ulem}   % temporal one in draft.
\usepackage{xspace}

%\usepackage{harvard}
\usepackage{amsopn} % for using DeclareMathOperator

@@

\DeclareMathOperator{\erf}{erf}

\begin{document}

@@

\author{L.~Wright$^1$, M.~Pitkin$^1$ \& C.~Messenger$^1$}
\address{$^1$ SUPA, School of Physics and Astronomy, University of
  Glasgow, Glasgow G12 8QQ, United Kingdom}
\eads{\mailto{matthew.pitkin@glasgow.ac.uk}, \mailto{christopher.messenger@glasgow.ac.uk}}

\begin{abstract}
  We present an investigation into the potential of calibrating \MP{change to: ``assessing the validity of the calibration''?}
  gravitational wave detector outputs through the use of standard
  sirens. Such signals, as measured via gravitational wave
  observations provide an estimated luminosity distance which is
  subject to uncertainties in the calibration of the data.  If a host
  galaxy is identified for a given source then its redshift can be
  obtained and combined with current knowledge of the cosmological
  parameters.  This will yield the true luminosity distance and allow
  the direct comparison with the estimated value.  Discrepancies can
  then be used to correct the original calibration.~\cm{Update at end}
\end{abstract}

% acronym definitions
\acrodef{GW}[GW]{gravitational wave}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{MWEG}[MWEG]{Milky Way equivalent galaxies}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{NSBH}[NSBH]{neutron star black hole}
\acrodef{SNR}[SNR]{signal-to-noise ratio}
\acrodef{GRB}[GRB]{gamma-ray burst}
\acrodef{sGRB}[sGRB]{short gamma-ray burst}
\acrodef{CBC}[CBC]{compact binary coalescence}
\acrodef{EM}[EM]{electro-magnetic}

@@

% introduce GWs and BNS systems
It is expected that the advanced generation of interferometric \ac{GW}
detectors will detect waves emitted from $O(10s)$ of compact binary
coalescences.  One such class of these cataclysmic events, the inspiral and
merger of \ac{BNS} systems will be detected out to a maximum range of $\approx
450$ Mpc. Assuming the current best estimates for the cosmological parameters,
this is equivalent to a redshift $z\approx 0.1$. As noted by \cite{1986Natur.323..310S} \ac{CBC}
systems can be used as cosmological distance markers, otherwise known as
``standard sirens'' (analogous to the \ac{EM} standard candles). Standard sirens
allow ...

@@

% Describe our model for the calibration
\MP{The gravitational wave strain is measured through the differential arm length, $\Delta L$, changes
of the interferometer via $h = \Delta L(f) / L$, where $L$ is the full arm length.
Calibration is required to relate the actual measured interferometer error signal output
$e(f)$ to $\Delta L$. This relation is known as the length response function, $R(f,t)$,
defined such that
\begin{equation}
\Delta L(f) = R(f,t) e(f),
\end{equation}
where $R$ varies slowly in time (in comparison to transient signal timescales).
Calculation of $R$ requires the measurement of various functions within a control feedback loop
(see \cite{Vitale:2012} for a brief overview of this), which are subject to
uncertainties. In this study we will assume an estimate of $R$ is available (although
in theory we could take on the role of estimating $R$ itself), but that it differs from the
truth through some unknown scale factor, $s$, so that
\begin{equation}
h(f) = s h_{\rm m}(f) = s \frac{R(f,t) e(f)}{L},
\end{equation}
where $h(f)$ is the true strain and $h_{\rm m}(f)$ is the measured strain.
In this analysis we will simplify the situation by assuming that $s$ is a constant with
respect to frequency, but in future studies that assumption could be dropped and $s$
could take some functional form, or piece-wise fit, with respect to $f$.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysis}

@@

The frequency-domain polarisation amplitudes are then defined as
%
\begin{eqnarray}\label{eq:signal}
  \gws_{+}(f) &=& \frac{1+\cos^{2}\iota}{2D_{L}}
\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)} \nonumber \\
  \gws_{\times}(f) &=&
 i\frac{\cos^{2}\iota}{D_{L}}\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)}
\end{eqnarray}
%
where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$,
and $\iota$ is the inclination angle.  The \ac{GW} strain measured
at the $k$'th detector is then given by
%
\begin{equation}
  \label{eq:gravsig}
   \gws^{k}(f) = \big[ F_{+}^{k}(\alpha, \delta, \psi)\gws_{+}(f) +
F_{\times}^D(\alpha, \delta, \psi)\gws_{\times}(f)\big]
\end{equation}
%
where $F_{+},F_{\times}$ are the antenna response functions which are dependent
upon the polarisation angle $\psi$ and the sky position of the source
$\alpha,\delta$.

@@

% Define the likelihood function we will be using
We consider that the measured data in any detector is defined as
%
\begin{equation}
  \tilde{d}_{k}(f)=\tilde{n}_{k}(f)+s\gws_{k}(f,\btheta)
\end{equation}
%
where $\tilde{n}(f)$ is the Fourier transform of Gaussian noise drawn from a
distribution consistent with the power-spectral-density of the advanced
detector noise.

\subsection{Method}

\MP{Our aim is to estimate the marginal posterior probability distribution function (pdf)
for the unknown calibration scale factor. We use Bayes' theorem for which we need to define
a likelihood function and prior pdfs on the signal parameters being estimated. In this
analysis we use the Markov chain Monte Carlo code {\tt emcee} \cite{2013PASP..125..306F} to sample 
the posterior distribution.} We use a Gaussian likelihood function given by
%
%\begin{equation}
%  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta
%f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})}\Bigg]
%\end{equation}
\begin{equation}
\fl p(\bi{d} | \btheta, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm det}}
  \sum_{i=i_{f_{\rm low}}}^{i_{f_{\rm high}}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} 
\gws_{k,i}(\btheta) - \frac{1}{2}\left(\gws^*_{k,i}(\btheta)\gws_{k,i}(\btheta) + 
\tilde{d}^*_{k,i}\tilde{d}_{k,i}\right) \right\}}}{S_{k,i}}\right]}
\end{equation}
% 
where $\Delta f$ is the frequency resolution of the data, $S_k$ is the $k^{\rm th}$ detector's one 
sided noise power spectral density, and the $i$ indices increment over frequency with $i_{f_{\rm 
low}}$ and $i_{f_{\rm high}}$ corresponding to the lower and upper range in frequencies used for 
our analysis (which were 40\,Hz to 1600\,Hz respectively \MP{Should we actually go down to 20\,Hz -- 
I imagine for BNS systems it doesn't make much difference?}). The source parameters, and the 
unknown scale factor, have been subsumed into a vector $\vec{\theta}$. For a network of detectors, 
the overall likelihood function becomes a product of the likelihoods for each detectors, i.e.\
\begin{equation}
  \label{eq:mult-likeli}
  p(\bi{d}| \btheta, \curlH) = \prod \limits_D p(\bi{d}^D| \vec{\theta}, \curlH)
\end{equation}
where $D = \{H,L,V\}$.

@@

\subsubsection{Prior ranges}\label{sec:priors}

The primary reason why we can use coincident observations with a GRB as a potential check of 
detector calibration is that the GRB allows us to constrain our priors on various parameters that 
would normally be highly correlated with the signal amplitude. The most important of these is that 
the GRB can provide a very tight constraint on the source distance\footnote{A consequence of our 
choice of BNS signals is that they will only be observable within the local Universe ($z \sim 
0.1$), so the relation between the measured GRB redshift and distance is simple and is negligably 
effected by the choice of cosmology.}. In this analysis we therefore assume that the source distance 
is known (i.e.\ has a $\delta$-function prior). Another piece of information that we can use from 
having a coincident GRB is that the system must be close to face-on, which allows us to place prior 
constraints on the system inclination angle. In our analysis we take two scenarios: the first use a 
prior on the beam opening angle (half-width of the beam) with a Gaussian distribution with zero mean 
and $\sigma = 5^{\circ}$, whilst the second uses $\sigma = 20^{\circ}$. Beam opening angles for 
short GRBs are relatively poorly constrained as they are based on only a few jet break detection, 
however \cite{2014ApJ...780..118F} give a median opening angle value of $\sim 10^{\circ}$. If we 
assume that the beam opening angle is a half-normal distribution, for which the median is given by 
$\sqrt{2}\erf{}^{-1}(1/2)\sigma \sim 0.67\sigma$, then a median of $\sim 10^{\circ}$ corresponds to 
$\sigma \sim 14.8^{\circ}$.

As we have assumed that the GRB progenitor is a binary neutron star system the prior we use for the 
system component masses is a Gaussian distribution for both components with means 
of $1.35\,\textrm{M}_{\odot}$ and standard deviations of $0.13\,\textrm{M}_{\odot}$ 
\MP{(reference)}. In reality, for all the systems we use in our study, the chirp mass, which 
we see from equation~\ref{eq:signal} plays a role in the overall signal amplitude, is very well 
constrained by the systems phase evolution, so our prior could be expanded with minimal effect on 
the results. In the future neutron star-black hole systems could be included, although this add the 
complication of having to include the black hole spin (by using BNS systems we have assumed that 
the spins are small enough that they will introduce only a negligible effect).

The final important piece of information that we can make use of from the GRB observation is the 
sky position of the source, which we take to be known precisely.

For the time of coalescence we use a uniform prior of spanning $\pm0.01$\,seconds around the 
recorded time of the signal that would be returned by the detection pipeline. For the reference 
phase a uniform prior between 0 and $2\pi$ is used, and for the polarisation angle a uniform prior 
between 0 and $\pi/2$ is used. As the signals we use are generally close to being circularly 
polarised (face-on) the phase and polarisation angle will be degenerate meaning the exact 
combination of them will have very little effect of the result.

Finally, we require a prior on the calibration scale factors for each detector. We make the 
assumption that in general the calibration will be correct, so want to use a prior that is peaked 
at unity. We also assume that the probability density that the calibration scale factor is either 
larger or smaller than unity by an equivalent factor, e.g.\ $s=10$ or $s=0.1$, is the same. For this 
we therefore use a log-normal distribution as our prior
\begin{equation}
 p(s|I) = \frac{1}{s\sigma\sqrt{2\pi}}\exp{\left( -\frac{(\ln{s} - \mu)^2}{2\sigma^2} \right)},
\end{equation}
where we chose $\sigma = 1$, which in turn means that having the mode at unity gives $\mu = 1$. 
With these parameters the prior probability density at 0.1 and 10 is $\sim 1/14$ of that at 1. 
Note that this prior does not provide produce a symmetric amount of probability about unity, but 
given our simulation criterion described below in Section~\ref{sec:simulations} we find that the 
likelihood generally overwhelms the prior. As $s$ is a scale factor we also have the further 
constraint that it must be positive.

\subsection{Simulations}\label{sec:simulations}

To estimate how well we can constrain the calibration scaling each detector in an advanced 
detector network we have performed simulations with injected BNS signals at a range of distances 
(50\,Mpc to 450\,Mpc on 50\,Mpc increments). The simulations use the two Advanced LIGO detectors 
(one at Hanford -- H1 -- and one at Livingston -- L1) and the Advanced Virgo detector (V1). We 
assume all are expected to be operating at their design sensitivity.

% Describe the ensemble of simulations used
As described in Section~\ref{sec:priors} we had two different scenarios for the prior distribution 
of source system inclination angles, so we have produced separate simulations for both cases. In 
each case 1000 sources were created with distributions on the inclination, component masses,
reference phase and polarisation angle drawn from their respective prior distributions. For the 
time of coalescence the times were held fixed within the centre of the prior window. When drawing 
the sky position points were drawn uniformly over the sky. However, with all these an additional 
constraint was imposed based on the fact that we require a distributions sources that could 
actually have been detected. We therefore include the requirement from \cite{2012PhRvD..85h2002A} 
that their be at least two detectors in the network for which the SNR is $\gtrsim 5.5$ (as these 
would be signals coincident with a GRB we do not include the often used further constraint that 
the total coherent network SNR is greater than 12). This generally means that as we go to larger
distances our population of sources that fulfil this criterion will not be uniformly distributed
over the sky, and will generally be closer to being circularly polarised.

For each source we applied a fixed calibration scale factor of 1.2, 0.8 and 1.3 for H1, L1 and V1 
respectively.

These simulations allow us to assess how well on average we would be able to constrain the 
calibration scale for a given CBC-GRB coincidence.

\section{Results}\label{sec:results}

For each of the simulated sources we have calculated the 68\% credible region on the scale factors 
for each detector (for a Gaussian distribution this is equivalent to the region bounded by 
$1\sigma$ intervals). For all the signals in each distance increment we have produced the 
distribution of the half widths (i.e.\ $1\sigma$) of these scale factors. These are shown in 
figures~\ref{fig:narrowdist} and \ref{fig:widedist} for the case with the inclination distribution 
with a width of $5^{\circ}$ and $20^{\circ}$ respectively.

\begin{figure}
 \begin{center}
  \includegraphics[width=1.0\textwidth]{scale_factors_5degs_boxplot.png}
 \end{center}
 \caption{\label{fig:narrowdist} Add caption!}
\end{figure}

\begin{figure}
 \begin{center}
  \includegraphics[width=1.0\textwidth]{scale_factors_20degs_boxplot.png}
 \end{center}
 \caption{\label{fig:widedist} Add caption!}
\end{figure}
